---
title: "Session 02 – Vietnamn GDP (World Bank WDI)"
output:
  html_document:
    toc: true
    toc_depth: 2
    number_sections: true
    theme: flatly
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
date: "`r Sys.Date()`"
---

## GDP Indicator Overview

In this session we use World Bank WDI indicators for Vietnam (`VNM`), including:

- **NY.GDP.MKTP.KD** – GDP (constant 2015 US$)  
- **NY.GDP.MKTP.KD.ZG** – GDP growth (annual %)

These will be used to illustrate level and growth of real GDP over time.

## World Bank Dashboard

You can explore GDP visually using the World Bank dashboard:

[World Bank – GDP Growth World Map](https://databank.worldbank.org/embed/GDP-Growth-Map/id/c8ac38a2)

## Loading Packages and Downloading Data

```{r setup_VNM, message=FALSE, warning=FALSE}
# install.packages("WDI") # run once in the Console if not installed
library(WDI)
```

```{r download_gdp_VNM}
# Download the full time series of Vietnamn GDP:
#   - NY.GDP.MKTP.KD    : GDP (constant 2015 US$)
#   - NY.GDP.MKTP.KD.ZG : GDP growth (annual %)
gdp_VNM <- WDI(
  country   = "VNM",
  indicator = c("NY.GDP.MKTP.KD", "NY.GDP.MKTP.KD.ZG")
)

# Sort from earliest to latest year
gdp_VNM <- gdp_VNM[order(gdp_VNM$year), ]

# Preview the first and last observations
head(gdp_VNM)
tail(gdp_VNM)
```

## Checking the Data Structure

```{r structure_gdp_VNM}
str(gdp_VNM)
```

## Searching for GDP-Related Indicators

```{r search_indicators_VNM, message=FALSE}
# Example: search WDI for indicators containing the word "growth"
# ind_growth <- WDIsearch("growth")
# head(ind_growth, 100)
```

## Plotting Vietnamn GDP (constant 2015 US$)

```{r plot_gdp_level_VNM, echo=FALSE}
plot(
  gdp_VNM$year,
  gdp_VNM$NY.GDP.MKTP.KD,
  type = "l",
  xlab = "Year",
  ylab = "GDP (constant 2015 US$)",
  main = "Vietnamn GDP (constant 2015 US$)"
)
```

## Plotting Vietnamn GDP Growth (annual %)

```{r plot_gdp_growth_VNM, echo=FALSE}
plot(
  gdp_VNM$year,
  gdp_VNM$NY.GDP.MKTP.KD.ZG,
  type = "l",
  xlab = "Year",
  ylab = "GDP growth (annual %)",
  main = "Vietnamn GDP Growth (annual %)"
)
```

